Deriving Hamiltonian mechanics from physics, without passing through Lagrangians.

### 2014-06-15 2014ClassicalHamiltonianDerivation/classicalHamiltonianDerivation.tex
Reviewed explanation for time d.o.f.

\begin{prop}\label{sdofMap}
Let $f: X \mapsto X$ be a bijective map on a degree of freedom. Then $dx' \wedge dp' = dx \wedge dp$.
\end{prop}

This is the generalization of \ref{widthMapping} for two state variable. The density $\rho(p,q)$ will be defined on cells of infinitesimal area proportional to $dx \wedge dp$. When mapping one cell to another, the infinitesimal area will remain the same. The area in phase space of one degree of freedom can then be used as a measure for the cardinality of the labels. \ref{labelsCount} becomes area conservation for conjugate variables. Area conservation is equivalent to requiring the invariance of the vector product.

@@

The orthogonality of the labels corresponds to orthogonality in phase space, which the mapping needs to preserve across independent degrees of freedom. The cardinality of labels within each degree of freedom corresponds to the area in phase space\footnote{We assume we are using the same unit across d.o.f.}, which the mapping also needs to preserve. This is equivalent to require the conservation of the scalar product across independent degrees of freedoms, while still requiring conservation of the vector product within. That leads us to the metric defined by \ref{mdofInvariance}.
The metric generalizes \ref{sdofInvariance} to give us the cardinality of labels defined on the area given by two arbitrary vectors. For an infinitesimal region, this corresponds to $dx^1 \wedge dp^1 + dx^2 \wedge dp^2$, the sum of the projections on the independent planes.

@@

\begin{prop}\label{tdofInvariant}
Let $v$ and $w$ be two vectors defined on the tangent space of the extended phase space for one degree of freedom. Let $\alpha$ and $\beta$ be indexes for $t, E, x, p$. Let
\begin{align*}
\omega_{\alpha, \beta} = \omega_{\alpha, \beta} =  \left[
  \begin{array}{cc}
    -1 & 0 \\
    0 & 1 \\
  \end{array}
\right] \otimes \left[
  \begin{array}{cc}
    0 & 1 \\
    -1 & 0 \\
  \end{array}
\right]
= \left[
  \begin{array}{cccc}
    0 & -1 & 0 & 0 \\
    1 & 0 & 0 & 0 \\
    0 & 0 & 0 & 1 \\
    0 & 0 & -1 & 0 \\
  \end{array}
\right] \\
\end{align*}
then $v'^{\alpha} \omega_{\alpha, \beta} w'^{\beta}=v^{\alpha} \omega_{\alpha, \beta} w^{\beta}$ under detrev transformation.
\end{prop}

$\langle T, E \rangle$ are not independent labels from $\langle Q, P \rangle$ as they do not define new states. They are not necessarily orthogonal directions in the extended phase space. Looking back at \ref{disclabelsoverr}, cells need to be defined on the plane where $\langle Q, P \rangle$ changes: they are not defined on the plane of constant $\langle E, T \rangle$ (they are not orthogonal), but on the plane perpendicular to $\langle Q, P \rangle$. Areas defined on that plane are then going to be invariant.

Like before, we have three planes of conjugate variables forming a right triangle-like relationship with one side invariant. In this case, the right angle is between the invariant and the plane of constant $\langle Q, P \rangle$ (on which $dt$ and $dE$ are defined). So we must have $dt \wedge dE + k = dq \wedge dp$. Which corresponds to the Minkowski product across d.o.f. and the vector product within. The metric, with a space-like convention, gives us a measure for states.

>>>>

\begin{prop}\label{sdofMap}
Let $f: X \mapsto X$ be a bijective map on a degree of freedom. Then $dq' \wedge dp' = dq \wedge dp$.
\end{prop}

This is the generalization of \ref{widthMapping} for two state variable. The density $\rho(p,q)$ will be defined on cells of infinitesimal area proportional to $dq \wedge dp$. When mapping one cell to another, the infinitesimal area will remain the same. The area in phase space of one degree of freedom can then be used as a measure for the cardinality of the labels. \ref{labelsCount} becomes area conservation for conjugate variables. Area conservation is equivalent to requiring the invariance of the vector product.

@@

The orthogonality of the labels corresponds to orthogonality in phase space, which the mapping needs to preserve across independent degrees of freedom. The cardinality of labels within each degree of freedom corresponds to the area in phase space\footnote{We assume we are using the same unit across d.o.f.}, which the mapping also needs to preserve. This is equivalent to require the conservation of the scalar product across independent degrees of freedoms, while still requiring conservation of the vector product within. That leads us to the metric defined by \ref{mdofInvariance}.
The metric generalizes \ref{sdofInvariance} to give us the cardinality of labels defined on the area given by two arbitrary vectors. For an infinitesimal region, this corresponds to $dq^1 \wedge dp^1 + dq^2 \wedge dp^2$, the sum of the projections on the independent planes.

@@

\begin{prop}\label{tdofInvariant}
Let $v$ and $w$ be two vectors defined on the tangent space of the extended phase space for one degree of freedom. Let $\alpha$ and $\beta$ be indexes for $t, E, q, p$. Let
\begin{align*}
\omega_{\alpha, \beta} =  \left[
  \begin{array}{cc}
    -1 & 0 \\
    0 & 1 \\
  \end{array}
\right] \otimes \left[
  \begin{array}{cc}
    0 & 1 \\
    -1 & 0 \\
  \end{array}
\right]
= \left[
  \begin{array}{cccc}
    0 & -1 & 0 & 0 \\
    1 & 0 & 0 & 0 \\
    0 & 0 & 0 & 1 \\
    0 & 0 & -1 & 0 \\
  \end{array}
\right] \\
\end{align*}
then $v'^{\alpha} \omega_{\alpha, \beta} w'^{\beta}=v^{\alpha} \omega_{\alpha, \beta} w^{\beta}$ under detrev transformation.
\end{prop}

$\langle T, E \rangle$ are not independent labels from $\langle Q, P \rangle$ as they do not define new states. So they are not necessarily orthogonal in the extended phase space. Looking back at \ref{disclabelsoverr}, cells need to be defined on the plane where $\langle Q, P \rangle$ (maximally) changes: this is not defined on the plane of constant $\langle E, T \rangle$ (they are not orthogonal) where $dq \wedge dp$ is defined, but on the plane perpendicular to constant $\langle Q, P \rangle$ where $dt \wedge dE$ is defined. On that plane we can properly count states and define our invariant.

As in the case of independent d.o.f., we have three planes of conjugate variables forming a right triangle-like relationship with one side invariant.
\begin{align*}
dq^1 \wedge dp^1 + dq^2 \wedge dp^2 = k \\
dt \wedge dE + k = dq \wedge dp
\end{align*}
In the previous case, the right angle was between the two independent degree of freedom. In this case, the right angle is between the invariant and the plane of constant $\langle Q, P \rangle$ where $dt \wedge dE$ is defined. We rewrite it as $dq \wedge dp - dt \wedge dE = k$. Which corresponds to the Minkowski product across d.o.f. and the vector product within. The metric, with a space-like convention, still gives us the cardinality of labels within a degree of freedom.
